Quality Assesment of OpenStreetMap Footpath Data for 3D City Modelling

### 2020-05-07 report/report.tex
Restructure introduction
Rename OpenStreetMap -> OSM

This project assesses the feasibility of creating a 3D representation of OpenStreetMap (OSM) geodata with the inclusion of pedestrian footpaths, by focusing on OSM road and building data in the Stockholm metropolitan area.
\\
An estimate of the positional accuracy of OSM data has been obtained by comparing OSM positional data to a dataset with a known accuracy, which was obtained from the Swedish National Land Survey.
It was found that the average positional error in the OSM dataset was about 2 meters, and approximately 98\% of all buildings in the reference dataset were represented in the OSM dataset.
\\
This estimate of the positional error was used to assess the integrity of the OSM Stockholm road network, by locating critical points where extending the roads and paths with indivdual widths would cause collision with other features.
An algortitm then took the road network and made adjustments that were within the found margin of error for the OSM dataset, to eliminate as many collisions as possible.
It was found that extending all roads with a standard width led to collisions between roughly 12\% of all roads in the dataset. The total length of road that intersected with other features was less than 0.1\% of the total road length. It was measured that roughly 80\% of these colliding features could be corrected by simple geometrical translation by a distance smaller than the posiional accuracy of the OSM dataset. Similarily, the amount of colliding road by length that could be corrected was calculated to around 99\%.
\\
The report concludes with a discussion of the findings of this study, and an assessment of the feasibility of generating city models with pedestrian footpaths in a number of different applications from urban planning tools to videogames.
As secondary contributions, the completeness and shape accuracy of the OSM data set have also been estimated and commented.

@@

This report covers... \cite{sampleitem2}

@@

This project has dealt with the feasibility of including footpath data when generating 3D representations of OpenStreetMap data.

The primary contribution is an investigation into the rate of collision between map features in OpenStreetMap. In order to create a 3D representation of geodata, individual features will be extruded to give them width, which may create situations where meshes are intersecting. The question is: how can a program identify critical areas where the footpath width is so large that it collides with existing roads? Can these collisions be easily fixed by making adjustments that are within the positional accuracy of the OpenStreetMap data? The features will be categorized according to type, so that the rate of collisions between footpaths and other features can be compared to that of other way types (primary, secondary, residential).

The secondary contribution will be an assesment of the geometrical precision of OpenStreetMap data around the Stockholm area. The positional accuracy will be needed to assess where road collisions occur and whether they can easily be corrected.

With that said, Nishida et. al. present an excellent solution for generating new road patches that comform to an underlying terrain function.

@@

In recent years, there have been an advent of tools that extract geodata from OpenStreetMap (OSM) and generate 3D representations of real world locations by placing terrain, buildings, roads and other features.
However, freely available such tools seldom include smaller roads and pedestrian footpaths such as walkways, sidewalks or trails.

@@

Since then a number of publications have examined slices of geodata from van Oort's criteria. In 2008, (Haklay, 2010) conducted such a study to estimate of the quality of OpenStreetMap positional geodata in London and England. Since the OpenStreetMap (OSM) project started in London it was thought that OSM data in the London metropolitan area would be representative of the highest quality data available, and therefore a good indicator for the whole global OSM dataset. The study was conducted by comparing OSM data to Ordnance Survey datasets. The study showed that the OSM dataset had a rough positional accuracy of 6 meters from the reference dataset. However, at the time of the study, the OSM project had captured roughly 29\% of the area of England, meaning that the data completeness was fairly low.

The next notable study to assess the quality of OpenStreetMap data was published by (Kunze, 2012). The study applied different methods to assess the completeness of OSM data in two federal states in Germany, mainly by analysing the area difference between the OSM dataset and an administrative dataset.

@@

For the purpose of this project, the OpenStreetMap road dataset has been used in the presentation of the road network around Stockholm, and in the generative process of new road networks. To assess and correct path widths and path node locations, the OpenStreetMap building footprint dataset has also been used. To make an assessment on the quality and precision of the OpenStreetMap data, the SLU map from the Swedish National Land Survey was used in comparison with the OpenStreetMap map. The SLU map conatins only building footprints and property limits. Both map segments were obtained around the Stockholm greater metropolitan area.
The exact map segments that were extracted were in both cases rectangular boundaries with the following coordinates in WGS84: (N: 59.42, E: 18.15, S: 59.23, W: 17.79).
Both datasets were collcted on March 18th, 2020.

@@

Volunteered Geographical Information.
OpenStreetMap road data and building footprints are commonly obtained by manually tracing features in commercially available satellite images. Such images have a limit on their resolution which puts a theoretical limit on the accuracy of map features compared to their real-world equivalents (Haklay 2010). Particularily the high resolution imagery from Bing in 2010 led to an increase in building information in OpenStreetMap (Fan et al, 2014).

@@

The SLU dataset is delivered in the SWEREF 99 TM (Lantmäteriet, 2019). SWEREF 99 TM is a projected coordinate system and there is no linear transformation to WGS 48, which OpenStreetMap uses (OSM 19). The coordinate conversions for this paper were obtained using proj: a Linux commandline application for geospatial coordinate conversion.

@@

(Fan et al., 2014) used the total building area in their reference dataset and the OpenStreetMap dataset to assess the completeness of OpenStreetMap data. The motivation for this is to eliminate semantic differences between both datasets. A very common example of a semantic error when comparing geographic data is that a large building in one dataset may be segmented into several smaller ones in the other, that is to say a building in one set is represented as an aggreation of mutliple buildings in the other set. This makes one-to-one object mapping impossible. Thus using building area as an estimator for completeness solves this issue much better than i.e. the number of buildings or other objects in each dataset.
This project will use the building area in the OSM dataset versus that of the SLU dataset to determine the completeness of OpenStreetMap building data in the Stockholm area, which will then be used as an argument for how accurate the building correspondence and point proximity measures can be.

@@

(Fan et al., 2014) further discussed how the relative overlap in building area can be used to determine building correspondence between datasets, in cases where there is not much displacement between OSM building footprints data and the reference data set. Figure X (TODO) shows a typical comparison between the OpenStreetMap and SLU datasets, and by eyesight it was determined that the displacement of features is small. The relative building overlap between the OSM and SLU datasets will be defined as folows, given the footprints of any building in the OSM set ($foot_{OSM}$) and the SLU set ($foot_{SLU}$):

@@

Finally, when two building footprints and their vertices have been mapped, we will use the offset of matching vertices as a measure of positional error. The average and variance in vertex spread will be calculated and used to make an assessment of the geometric precision of the OpenStreetMap dataset, and this precision will further be used to form an upper boundary on geometrical corrections that can be made to the map once it has been imported into the application.

@@

Accuracy: we can assume that Stockholm, being one of the more developed cities in one of the most technologically connected country in the world, would be fairly well mapped in OpenStreetMap. The positional accuracy has improved by roughly 3.5 meters since the study by (Haklay, 2010), and by roughly 1.5 meters since the study by (Fan et al, 2014), if the SLU reference dataset is to be believed as absolute.

@@

The difference in area cover can partially be explained by small utility buildings that are often not represented in the OpenStreetMap dataset. Particularily the residential areas in downtown, Stockholm, are characteristic in that each city block consists of a street front and an open, communal compound in the middle, inside of which there are usually a number of small utility buildings such as bicycle garages, toolsheds, refuse rooms and such. These buildings are likely hard to make out and trace on publicly available satellite image data. They are, according to (Fan et al, 2014) occluded by their surroundings due to shadows or forestry. This appears to be a large consensus in the field.

>>>>

This project assesses the feasibility of creating a 3D representation of OpenStreetMap (OSM) geodata with the inclusion of pedestrian footpaths, by focusing on OSM road and building data in the Stockholm metropolitan area.
\\
An estimate of the positional accuracy of OSM data has been obtained by comparing OSM positional data to a dataset with a known accuracy, which was obtained from the Swedish National Land Survey.
It was found that the average positional error in the OSM dataset was about 2 meters, and approximately 98\% of all buildings in the reference dataset were represented in the OSM dataset.
\\
This estimate of the positional error was used to assess the integrity of the OSM Stockholm road network, by locating critical points where extending the roads and paths with indivdual widths would cause collision with other features.
An algortitm then took the road network and made adjustments that were within the found margin of error for the OSM dataset, to eliminate as many collisions as possible.
It was found that extending all roads with a standard width led to collisions between roughly 12\% of all roads in the dataset.
The total length of road that intersected with other features was less than 0.1\% of the total road length.
It was measured that roughly 80\% of these colliding features could be corrected by simple geometrical translation by a distance smaller than the posiional accuracy of the OSM dataset.
Similarily, the amount of colliding road by length that could be corrected was calculated to around 99\%.
It was found that footpaths were much less problematic than other road types (primary, secondary, residential), in terms of feature collisions.
\\
The report concludes with a discussion of the findings of this study, and an assessment of the feasibility of generating city models with pedestrian footpaths in a number of different applications from urban planning tools to videogames.
As secondary contributions, the completeness and shape accuracy of the OSM data set have also been estimated and commented.

@@

@@

This project has dealt with the feasibility of including footpath data when generating 3D representations of OpenStreetMap (OSM) data.
\\
The primary contribution is an investigation into the rate of collision between map features in OSM.
In order to create a 3D representation of geodata, individual features will be extruded to give them width, which may create situations where meshes are intersecting.
The question is: how can a program identify critical areas where the footpath width is so large that it collides with existing roads?
Can these collisions be easily fixed by making adjustments that are within the positional accuracy of the OSM data?
The features will be categorized according to type, so that the rate of collisions between footpaths and other features can be compared to that of other way types (primary, secondary, residential).
\\
The secondary contribution will be an assesment of the geometrical precision of OSM data around the Stockholm area.
The positional accuracy will be needed to assess where road collisions occur and whether they can easily be corrected.

@@

In recent years, there have been an advent of tools that extract geodata from OSM and generate 3D representations of real world locations by placing terrain, buildings, roads and other features.
However, freely available such tools seldom include smaller roads and pedestrian footpaths such as walkways, sidewalks or trails.

@@

Since then a number of publications have examined slices of geodata from van Oort's criteria. In 2008, (Haklay, 2010) conducted such a study to estimate of the quality of OSM positional geodata in London and England. Since the OSM project started in London it was thought that OSM data in the London metropolitan area would be representative of the highest quality data available, and therefore a good indicator for the whole global OSM dataset. The study was conducted by comparing OSM data to Ordnance Survey datasets. The study showed that the OSM dataset had a rough positional accuracy of 6 meters from the reference dataset. However, at the time of the study, the OSM project had captured roughly 29\% of the area of England, meaning that the data completeness was fairly low.

The next notable study to assess the quality of OSM data was published by (Kunze, 2012). The study applied different methods to assess the completeness of OSM data in two federal states in Germany, mainly by analysing the area difference between the OSM dataset and an administrative dataset.

@@

For the purpose of this project, the OOSMroad dataset has been used in the presentation of the road network around Stockholm, and in the generative process of new road networks. To assess and correct path widths and path node locations, the OSM building footprint dataset has also been used. To make an assessment on the quality and precision of the OSM data, the SLU map from the Swedish National Land Survey was used in comparison with the OSM map. The SLU map conatins only building footprints and property limits. Both map segments were obtained around the Stockholm greater metropolitan area.
The exact map segments that were extracted were in both cases rectangular boundaries with the following coordinates in WGS84: (N: 59.42, E: 18.15, S: 59.23, W: 17.79).
Both datasets were collcted on March 18th, 2020.

@@

Volunteered Geographical Information.
OSM road data and building footprints are commonly obtained by manually tracing features in commercially available satellite images. Such images have a limit on their resolution which puts a theoretical limit on the accuracy of map features compared to their real-world equivalents (Haklay 2010). Particularily the high resolution imagery from Bing in 2010 led to an increase in building information in OSM (Fan et al, 2014).

@@

The SLU dataset is delivered in the SWEREF 99 TM (Lantmäteriet, 2019). SWEREF 99 TM is a projected coordinate system and there is no linear transformation to WGS 48, which OSM uses (OSM 19). The coordinate conversions for this paper were obtained using proj: a Linux commandline application for geospatial coordinate conversion.

@@

(Fan et al., 2014) used the total building area in their reference dataset and the OSM dataset to assess the completeness of OSM data. The motivation for this is to eliminate semantic differences between both datasets. A very common example of a semantic error when comparing geographic data is that a large building in one dataset may be segmented into several smaller ones in the other, that is to say a building in one set is represented as an aggreation of mutliple buildings in the other set. This makes one-to-one object mapping impossible. Thus using building area as an estimator for completeness solves this issue much better than i.e. the number of buildings or other objects in each dataset.
This project will use the building area in the OSM dataset versus that of the SLU dataset to determine the completeness of OSM building data in the Stockholm area, which will then be used as an argument for how accurate the building correspondence and point proximity measures can be.

@@

(Fan et al., 2014) further discussed how the relative overlap in building area can be used to determine building correspondence between datasets, in cases where there is not much displacement between OSM building footprints data and the reference data set. Figure X (TODO) shows a typical comparison between the OSM and SLU datasets, and by eyesight it was determined that the displacement of features is small. The relative building overlap between the OSM and SLU datasets will be defined as folows, given the footprints of any building in the OSM set ($foot_{OSM}$) and the SLU set ($foot_{SLU}$):

@@

Finally, when two building footprints and their vertices have been mapped, we will use the offset of matching vertices as a measure of positional error. The average and variance in vertex spread will be calculated and used to make an assessment of the geometric precision of the OSM dataset, and this precision will further be used to form an upper boundary on geometrical corrections that can be made to the map once it has been imported into the application.

@@

Accuracy: we can assume that Stockholm, being one of the more developed cities in one of the most technologically connected country in the world, would be fairly well mapped in OSM. The positional accuracy has improved by roughly 3.5 meters since the study by (Haklay, 2010), and by roughly 1.5 meters since the study by (Fan et al, 2014), if the SLU reference dataset is to be believed as absolute.

@@

The difference in area cover can partially be explained by small utility buildings that are often not represented in the OSM dataset. Particularily the residential areas in downtown, Stockholm, are characteristic in that each city block consists of a street front and an open, communal compound in the middle, inside of which there are usually a number of small utility buildings such as bicycle garages, toolsheds, refuse rooms and such. These buildings are likely hard to make out and trace on publicly available satellite image data. They are, according to (Fan et al, 2014) occluded by their surroundings due to shadows or forestry. This appears to be a large consensus in the field.
